---
output: 
  github_document:
    toc: TRUE
    toc_depth: 2
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = T,
  error = T
)
```


# Part 00. Proposal

Proposing the {ggtextExtra} package! ðŸ¦„ 
<!-- (typical package introduction write up; but actually aspirational) -->

The goal of {ggtextExtra} is to make ... easier.


Without the package, we live in the effort-ful world that follows ðŸ‹:

```{r}
library(tidyverse)

iris  |>
 ggplot(aes(Sepal.Length, Sepal.Width,
             fill = Species)) +
  geom_point(show.legend = FALSE, size = 7, alpha = .8, shape = 21, color = "white") +
  scale_fill_viridis_d(end = .8) +
  # title color are manually added
  labs(title = "<span style = 'color: red;'>Virginica irises</span> have the largest average sepal width") +
  theme(plot.title = ggtext::element_markdown())

```

  
With the {xxxx} package, we'll live in a different world (ðŸ¦„ ðŸ¦„ ðŸ¦„) where the task is a snap ðŸ«°: 

Proposed API:

<!-- The following is fenced off to quote the code, but won't execute.  -->

```
my_title <- "The **setosa** irises has the largest average sapel widths <br>and then comes **virginica** irises while<br>**versicolor** has the shortest sapel width"

iris |>
 ggplot() +
  aes(x = Sepal.Length, 
      y = Sepal.Width,
      fill = Species) +
  geom_point(shape = 21, 
             color = "white", size = 8, alpha = .9) +
  scale_fill_viridis_d(end = .8, begin = .2) +
  labs(title = my_title) +
  theme(plot.title = ggtext::element_markdown())
  
use_fill_scale_in_title_words(plot = last_plot()) + 
  guides(fill = "none")  

```

# Part 0. Lay out package infrastructure

```{r, eval = F}
devtools::create(".")
```
# Part 00 Explore ggplot object internals


Now let's 'crack into ggplot2 internals' to see if we can get to a match between categories and colors programmatically.

First, we'll look at the colors actually rendered in the layer data of the plot using `layer_data`. These are often (always?) stored as hex colors which will definitely work in html/markdown context, whereas you have to be a little careful with some R named colors not working at all in html.

First, we'll use `layer_data` to grab fill and group.

```{r}
plot <- last_plot()

fill_values_df <- layer_data(plot) %>%  .[,c("fill", "group")] |> distinct()
fill_values_df
```

Then we'll grab the name of the variable that's mapped to fill. This seems a little weird, but seems to works. Open to different approaches that might be more robust!

```{r}
fill_var_name <- plot$mapping$fill |> capture.output() %>% .[2] %>% str_extract("\\^.+") %>% str_remove("\\^")
fill_var_name
```

Then we can grab the actual vector of data that's being represented by fill color - the `plot$data` slot. We put this in a dataframe/tibble, and then used `distinct` to get a one-to-one category-group table.

```{r}
fill_var <- plot$data %>% .[,fill_var_name] 
fill_var
fill_var_df <- tibble(fill_var, group = as.numeric(fill_var)) %>% distinct()
fill_var_df
```

Then we join our colors and categories by group, and have a color-category one-to-one table. And then we can prepare an html statement that will make the category colorful when rendered by ggtext functionality.

```{r}
left_join(fill_values_df, fill_var_df, by = join_by(group)) %>% 
  mutate(html_statements = paste0("<span style = 'color: ", fill, 
                         "'>", fill_var, "</span>") )

```


# Part I. Work out functionality  ðŸš§ âœ… 

Let's put this in a function. The function will let us go straight from a ggplot2 plot object to a dataframe that has the fill information.


```{r grab_fill_info}
grab_fill_info <- function(plot){
  
fill_values_df <- ggplot2::layer_data(plot) %>%  
  .[,c("fill", "group")] |> 
  dplyr::distinct()

fill_var_name <- plot$mapping$fill |> 
  capture.output() %>% .[2] %>% 
  stringr::str_extract("\\^.+") %>% 
  stringr::str_remove("\\^")

fill_var <- plot$data %>% 
  .[,fill_var_name] 

fill_var_df <- tibble::tibble(fill_var, group = as.numeric(fill_var)) %>%
  dplyr::distinct()


dplyr::left_join(fill_values_df, 
                 fill_var_df, 
                 by = dplyr::join_by(group)) %>% 
  dplyr::mutate(html_statements = 
                  paste0("<span style = 'color: ", 
                         .data$fill, 
                         "'>", 
                         .data$fill_var, 
                         "</span></strong>") )
  
}
```

We can test this out with the plot we saved before:

```{r}
grab_fill_info(plot = plot)
```

And then we'll use the data frame output, to make replacements in a string, adding the html tags. We'll just get it done with a for loop.  One danger, that I'm leaving for later, is that you might have a categories like 'anana' (this means pineapple in Portuguese and maybe some other languages) and 'banana' (this means banana in Portuguese and maybe some other languages).  In this case, you'll have a bad result given the current implementation. (Anana has an acent on the final syllable in Portuguese you might actually be saved!)  male/female is the same problem - but not as nice of a tangent.

```{r}
fill_df <- grab_fill_info(plot)

for(i in 1:nrow(fill_df)){

  start <- "virginica have the largest average sapel width"
  
  start <- start |> stringr::str_replace(fill_df$fill_var[i] %>% as.character(), fill_df$html_statements[i])

start %>% print()
  
}
```


Let's put the for loop in a function:

```{r auto_color_html}
auto_color_html <- function(x, fill_df ){
  
 for(i in 1:nrow(fill_df)){

  x <- x |> stringr::str_replace(fill_df$fill_var[i] %>% as.character, fill_df$html_statements[i])
  
 }
  
  x
  
}
```

Test it out...

```{r}
auto_color_html("The setosa iris is cool", grab_fill_info(plot))
```

# The big test of everything

Now let's use our functions with a fresh plot, q. 

```{r}
iris |>
 ggplot(aes(Sepal.Length, Sepal.Width,
             fill = Species)) +
  geom_point(show.legend = FALSE, shape = 21, color = "white", size = 8) +
  scale_fill_viridis_d(end = .8, begin = .2) +
  labs(title = "The **setosa** iris has the smallest average sapel width<br>and the **virginica** irises have largest average sapel width<br>
       while **versicolor** is in-between") +
  theme(plot.title = ggtext::element_markdown())

q <- last_plot()

q_fill_df <- grab_fill_info(q)

colorful_title <- "The **setosa** iris has the *smallest* average sapel width<br>and the **virginica** irises have *largest* average sapel width<br> while **versicolor** is in-between" |> 
        auto_color_html(q_fill_df)

q + 
  labs(title = colorful_title)  #overwriting title
```

# Further bundling

Looks good.  What if we wrap everything, and just replace the title with an html color-tagged version. 

```{r use_fill_scale_in_title_words}
use_fill_scale_in_title_words <- function(plot){
  
  out <- plot
  plot_fill_df <- grab_fill_info(plot)
  
  out$labels$title <- out$labels$title |> 
        auto_color_html(plot_fill_df)

  return(out)
  
}
```


Try it out starting fresh.

```{r}
my_title <- "The **setosa** irises has the largest average sapel widths <br>and then comes **virginica** irises while<br>**versicolor** has the shortest sapel width"

iris |>
 ggplot() +
  aes(x = Sepal.Length, 
      y = Sepal.Width,
      fill = Species) +
  geom_point(shape = 21, 
             color = "white", size = 8, alpha = .9) +
  scale_fill_viridis_d(end = .8, begin = .2) +
  labs(title = my_title) +
  theme(plot.title = ggtext::element_markdown())

use_fill_scale_in_title_words(plot = last_plot()) + 
  guides(fill = "none")
```






## Notes:

- fill mapping should be globally declared
- data should be globally declared
- should make function have an index for layer from which to get the fill information
- naive string replacement issue



# Part II. Packaging and documentation  ðŸš§ âœ… 

## Phase 1. Minimal working package

To build a minimal working package, 1) we'll need to document dependencies 2) send functions to .R files in the R folder, 3) do a package check (this will A. Document our functions and B. this will help us identify problems - for example if we've failed to declare a dependency) and 4) install the package locally.  

Then we'll write up a quick advertisement for what our package is able to do in the 'traditional readme' section.  This gives us a chance to test out the package.

```{r, eval = F}
### Bit 2a: in the function(s) you wrote above make sure dependencies to functions using '::' syntax to pkg functions 
usethis::use_package("ggplot2") # Bit 2b: document dependencies, w hypothetical ggplot2
usethis::use_package("dplyr")
usethis::use_package("stringr")
usethis::use_package("tibble")
usethis::use_pipe()

```

```{r}
knitrExtra::chunk_names_get()
# Bit 3: send the code chunk with function to R folder
knitrExtra:::chunk_to_r(chunk_name = "grab_fill_info") 
knitrExtra:::chunk_to_r(chunk_name = "auto_color_html")
knitrExtra:::chunk_to_r(chunk_name = "use_fill_scale_in_title_words")

```


```{r, eval = F}
# Bit 4: document functions and check that package is minimally viable
devtools::check(pkg = ".")  

# Bit 5: install package locally
devtools::install(pkg = ".", upgrade = "never") 
```


### Bit 7. Write traditional README that uses built package (also serves as a test of build). ðŸš§ âœ… 

The goal of the {xxxx} package is to ...

Install package with:

```
remotes::install_github("GithubCoolUser/mypacakge")
```

Once functions are exported you can remove go to two colons, and when things are are really finalized, then go without colons (and rearrange your readme...)

```{r, eval = F}
library(ggtextExtra)
my_title <- "The **setosa** irises has the largest average sapel widths <br>and then comes **virginica** irises while<br>**versicolor** has the shortest sapel width"

iris |>
 ggplot() +
  aes(x = Sepal.Length, 
      y = Sepal.Width,
      fill = Species) +
  geom_point(shape = 21, 
             color = "white", size = 8, alpha = .9) +
  scale_fill_viridis_d(end = .8, begin = .2) +
  labs(title = my_title) +
  theme(plot.title = ggtext::element_markdown())

ggtextExtra:::use_fill_scale_in_title_words(plot = last_plot()) + 
  guides(fill = "none")
```

# Bit 8: Compile readme

# Bit 9: Push to github

# Bit 10: listen and iterate

## Phase 3: Settling and testing ðŸš§ âœ… 

### Bit A. Added a description and author information in the [DESCRIPTION file](https://r-pkgs.org/description.html) ðŸš§ âœ…

### Bit B. Added [roxygen skeleton](https://r-pkgs.org/man.html) for exported functions. ðŸš§ âœ…

### Bit D. Settle on [examples](https://r-pkgs.org/man.html#sec-man-examples).  Put them in the roxygen skeleton and readme. ðŸš§ âœ…

### Bit C. Chosen a [license](https://r-pkgs.org/license.html)? ðŸš§ âœ…

```{r, eval = F}
usethis::use_mit_license()
```

### Bit D. Use life-cycle badge

```{r, eval = F}
usethis::use_lifecycle_badge("experimental") 
```


### Bit E. Written formal [tests](https://r-pkgs.org/testing-basics.html) of functions and save to test that folders ðŸš§ âœ…




That would look like this...

```{r test_calc_times_two_works, eval = F}
library(testthat)

test_that("calc times 2 works", {
  expect_equal(times_two(4), 8)
  expect_equal(times_two(5), 10)
  
})
```


```{r, eval = F}
knitrExtra::chunk_to_tests_testthat("test_calc_times_two_works")
```




### Bit F. Check again. Addressed notes, warnings and errors. ðŸš§ âœ…


```{r, eval = F}
devtools::check(pkg = ".")
```

## Phase 4. Promote to wider audience...  ðŸš§ âœ… 

### Bit A. Package website built? ðŸš§ âœ…


### Bit B. Package website deployed? ðŸš§ âœ…

## Phase 5: Harden/commit: Submit to CRAN/RUniverse ðŸš§ âœ… 

# Appendix: Reports, Environment

## Description file complete?  ðŸš§ âœ…

```{r, eval = F}
readLines("DESCRIPTION")
```

## Environment  ðŸš§ âœ…

Here I just want to print the packages and the versions

```{r}
all <- sessionInfo() |> print() |> capture.output()
all[11:17]
```

## `devtools::check()` report

```{r, eval=F, error = T, results="hide", warning=F}
devtools::check(pkg = ".")
```

## Package directory file tree

```{r}
fs::dir_tree(recurse = T)
```



